Journey: A Study of Scalabilty of Memory Management in Linux

### 2017-05-02 paper.tex
Update todo
Fix typo

\begin{figure}[t]
    \includegraphics[width=\columnwidth]{figures/mmap_latency}
    \caption{Latency of memory operations (TODO: improve the scale to make log behavior more apparent)}
    \label{fig:mmap_latency}
\end{figure}

@@

\section{Conclusion}
Terabytes scale memory systems are poised to become commonplace in not too distant future.
Developers and researchers need tools to study the behavior of such systems even before the actual
deployments are avaialable. In this paper, we study the scalability of memory management 
system of Linux Kernel which we believe is the first step toward building such tools. In particular, 
we look at the scalability of \texttt{vm\_area\_struct},  
swapping subsystem, and the effect of hugepages across a wide range of varying parameters
and workloads. Our findings indicate that save for the fragmented mmmap pattern, 
\texttt{vm\_area\_struct} tends to be sclable in the general case. We further find that the use of 
transparent hugepages can lead to early onset of memory pressure for the random access pattern of applications. 
This can lead to excessive swapping which might outweigh any potential benefits enabled by the use of hugepages. 
We also quantify the degradation in performance due to the use of large swapping files and find that very large 
swapping files should be used with caution becuase they can lead to very poor performance due to the loss of disk 
locality.This aspact is more crucial in large scale memory system compared the the conventional systems.  
While priliminary, we believe that our study is a first step towards building tools 
to study large memory systems in a principled and informed manner.

>>>>

\begin{figure}[t]
    \includegraphics[width=\columnwidth]{figures/mmap_latency}
    \caption{Latency of memory operations. A running average with window of
    width 5000 operations is used to reduce jitter.}
    \label{fig:mmap_latency}
\end{figure}

@@

\section{Conclusion}
Terabytes scale memory systems are poised to become commonplace in not too distant future.
Developers and researchers need tools to study the behavior of such systems even before the actual
deployments are avaialable. In this paper, we study the scalability of memory management 
system of Linux Kernel which we believe is the first step toward building such tools. In particular, 
we look at the scalability of \texttt{vm\_area\_struct},  
swapping subsystem, and the effect of hugepages across a wide range of varying parameters
and workloads. Our findings indicate that save for the fragmented mmap pattern, 
\texttt{vm\_area\_struct} tends to be sclable in the general case. We further find that the use of 
transparent hugepages can lead to early onset of memory pressure for the random access pattern of applications. 
This can lead to excessive swapping which might outweigh any potential benefits enabled by the use of hugepages. 
We also quantify the degradation in performance due to the use of large swapping files and find that very large 
swapping files should be used with caution becuase they can lead to very poor performance due to the loss of disk 
locality.This aspact is more crucial in large scale memory system compared the the conventional systems.  
While priliminary, we believe that our study is a first step towards building tools 
to study large memory systems in a principled and informed manner.
